\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2018-02-05 write_up/mixfishsim_ecolmod.tex
Manuscript edits and VAST fit to commercial data

%%\title{A simulation framework for exploring spatio-temporal dynamics in mixed
%%	fisheries}

@@

To implement effective spatial measures to reduce discards 
a good understanding of spatio-temporal fishery dynamics is required. However,
traditional scientific advice is limited by a lack of highly resolved knowledge of population distribution,
movement and how fishers interact with different fish populations. This reflects that 
data on fish location at high temporal and spatial resolutions is
expensive and difficult to collect and therefore proxies inferred from either
scientific surveys or commercial catch data are often used to model
distributions, often with limited spatial and temporal resolution. \\

To understand how resolution impacts mixed fisheries inference, we develop a highly 
resolved spatio-temporal simulation model incorporating:
i) delay-difference population dynamics, ii) population movement using
Gaussian Random Fields to simulate patchy, hetergenously distributed
populations, and iii) fishery dynamics for multiple fleet characteristics based on
targetting via correlated random walk movement and learned behaviour. \\

We simulate 20 years of exploitation of the fish populations
and use the results from the fishing model to draw inference on the underlying
population structures. We compare this inference to i) a simulated fixed-site
sampling design commonly used for fisheries monitoring purposes, and ii) the
true underlying population structures input to the simulation, to establish the potential and limitations of
fishery-dependent data - an inherently biased sampling method due to fisher's
targeting- to provide a robust picture of spatio-temporal distributions. [We
simulate a closure based on areas defined from commercial catch data and assess
its effectiveness on reducing catches of a fish population.] \\

@@

Use of spatial management as a tool has been proposed as a method to reduce
discards. However, its implementation is hampered by lack of knowledge of fish
and fishery spatiotemporal \textcolor{red}{[use this spelling throughout]} dynamics and understanding of the scale at which
processes are important for management. Understanding the correct scale for
spatial management is crucial in order to implement measures at a resolution
that ensures effective management\cite{Dunn2016} while minimising economic
impact. For example, a scale that promotes species avoidance for vulnerable or
low quota species while allowing continuance of sustainable fisheries for
available quota species. \\

@@

Using our model we simulate 20 years of exploitation of the fish populations
and use the results from the fishing model to draw inference on the underlying
population structures.  We compare this inference to: i) a stratified fixed-site sampling survey 
design commonly used for fisheries monitoring
purposes \textcolor{red}{[not sure if this is the fisheries-independent survey?]} and ii) the underlying population structures input to the simulation.\\

@@

% -------------------------------------------------
% A last flourish which breaks all the rules
\draw [->,purple, dotted, thick, shorten >=1mm]
  (p3.south) -- ++(5mm,-3mm)  -- ++(87mm,0mm) 
  |- node [black, near end, yshift=1.75em, it]
    {} ($(p1.north) + (0,0.3)$);
% -------------------------------------------------
\end{tikzpicture}
\caption{Overview Schematic of simulation model \textcolor{red}{Make a table with all variables and symbols so that you can reference them in the caption. Expand the caption to give as much standalone info as possible, e.g., what's blue, what's green, arrow colours, etc.}}
\label{fig:overview}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

where $C$ is the summed catch from the fishing model across all fleets and
vessels for the population during the day, and $B$ the daily biomass for the
species. \textcolor{red}{[link $F$ to effort and catchability - as I think we have F as an emergent property of the fleets rather than something we solve for (I could be wrong though!)]}\\

@@

\begin{equation}
	Pr(B | A) = \frac{e^{-\lambda * d_{AB}} \cdot
		Hab_{B}^2}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		Hab^2}
\end{equation}

Where $d_{AB}$ is the euclidean distance between cell $A$ and cell $B$, and
$\lambda$ is a given rate of decay. \textcolor{red}{[this is great but do explain $H_{ab}$ and all variables immediately on presentation throughout.]}\\

@@

with $k$ the concentration parameter from the von mises distribution which we
correlate with the revenue so that $k = (Rev + 1 / RefRev) * max_{k}$,
where $max_{k}$ is the maximum concentration value, $k$, and RefRev is
parameterised as for $\beta_{3}$ in the step length function \textcolor{red}{[wonderful!]}.

@@

\begin{itemize}
	\item Spatial dynamics: e.g. Figure 14. showing the population movement
		across weeks, including a spawning period.
	\item Overall population trends: e.g. Figure 3, showing the population
		dynamics at play.
	\item Realised step function: e.g. Figure 13, showing how the movement
		and turning angle responds. This could be combined with Figure
		7 for a more complete visualisaiton.
	\item Catch composition and spatial clustering: e.g. Figure 10, showing
		the importance of spatial resolution.
	\item Some measure of temporal trends/changes in spatial composition -
		could pick a certain cell and show how the proportions change
		over time ? An example is Figure 15, though there is little
		variation here. I think this is because the static habitat with
		population movement is far more stable than the previous
		directly affected population distributions (where there was an
		advective-diffusive process from the SPDE). Might need to
		consider how to change this?  Could include a spatio-temporally
		changing habitat suitability covariate, e.g. temperature ?? \textcolor{red}{[temperature would be very interesting and cover varying spatial fields]}
	\item Comparison of population structure from i) commercial sampling,
		ii) fisheries independent survey, iii) real population. This
		should be some statistical measure...not sure best approach
		here.
	\item [If we perform the closure, some measure of cost/benefit based on
		different resolutions of data and data used (i.e. commercial,
		survey, real populations). This could be on the basis of
		closing x cells, giving cell references to the model so they
		cannot be fished. Need to consider whether this is also a
		seasonal closure.]
\end{itemize}

>>>>

%%\title{A simulation framework for exploring spatiotemporal dynamics in mixed
%%	fisheries}

@@

To implement effective spatial measures to reduce discards a good understanding
of spatiotemporal fishery dynamics is required. However, traditional
scientific advice is limited by a lack of highly resolved knowledge of
population distribution, movement and how fishers interact with different fish
populations. This reflects that data on fish location at high temporal and
spatial resolutions is expensive and difficult to collect and therefore proxies
inferred from either scientific surveys or commercial catch data are often used
to model distributions, often with limited spatial and temporal resolution. \\

To understand how resolution impacts mixed fisheries inference, we develop a
highly resolved spatiotemporal simulation model incorporating: i)
delay-difference population dynamics, ii) population movement using Gaussian
Random Fields to simulate patchy, hetergenously distributed populations, and
iii) fishery dynamics for multiple fleet characteristics based on targetting
via correlated random walk movement and learned behaviour. \\

We simulate 20 years of exploitation of the fish populations and use the
results from the fishing model to draw inference on the underlying population
structures. We compare this inference to i) a simulated fixed-site sampling
design commonly used for fisheries monitoring purposes, and ii) the true
underlying population structures input to the simulation, to establish the
potential and limitations of fishery-dependent data - an inherently biased
sampling method due to fisher's targeting- to provide a robust picture of
spatiotemporal distributions. [We simulate a closure based on areas defined
from commercial catch data and assess its effectiveness on reducing catches of
a fish population.] \\

@@

Use of spatial management as a tool has been proposed as a method to reduce
discards. However, its implementation is hampered by lack of knowledge of fish
and fishery spatiotemporal dynamics and understanding of the scale at which
processes are important for management. Understanding the correct scale for
spatial management is crucial in order to implement measures at a resolution
that ensures effective management\cite{Dunn2016} while minimising economic
impact. For example, a scale that promotes species avoidance for vulnerable or
low quota species while allowing continuance of sustainable fisheries for
available quota species. \\

@@

Using our model we simulate 20 years of exploitation of the fish populations
and use the results from the fishing model to draw inference on the underlying
population structures.  We compare this inference to: i) a stratified
fixed-site sampling survey design commonly used for fisheries monitoring
purposes, otherwise know as a fisheries-independent
	survey, and ii) the underlying population structures input to the
simulation.\\

@@

% -------------------------------------------------
% A last flourish which breaks all the rules
\draw [->,purple, dotted, thick, shorten >=1mm]
  (p3.south) -- ++(5mm,-3mm)  -- ++(87mm,0mm) 
  |- node [black, near end, yshift=1.75em, it]
    {} ($(p1.north) + (0,0.3)$);
% -------------------------------------------------
\end{tikzpicture}
\caption{Overview Schematic of simulation model. The blue boxes indicate fleet
	dynamics processes, the green boxes population dynamics processes while
	the white boxes are the timesteps at which processes occur; $t$ = tow,
	$tmax$ is the total number of tows, $y$ = year, $pop_{t}$ is time of
	population movement, $day$ is a day timestep, $trip$ is a trip time
	step. 
	} 
\label{fig:overview}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

where $C$ is the summed catch from the fishing model across all fleets and
vessels for the population during the day, and $B$ the daily biomass for the
species. \textcolor{red}{[link $F$ to effort and catchability - as I think we
	have F as an emergent property of the fleets rather than something we
	solve for (I could be wrong though!) - catch for a vessel is a product of
	catchability and biomass, i.e. C = qB, but this catch is summed to
	solve for F. So its both really]}\\

@@

\begin{equation}
	Pr(B | A) = \frac{e^{-\lambda * d_{AB}} \cdot
		Hab_{B}^2}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		Hab_{B}^2}
\end{equation}

Where $d_{AB}$ is the euclidean distance between cell $A$ and cell $B$,
$\lambda$ is a given rate of decay and $Hab_{B}^2$ is the squared index of
habitat suitability for cell $B$.\\

@@

with $k$ the concentration parameter from the von mises distribution which we
correlate with the revenue so that $k = (Rev + 1 / RefRev) * max_{k}$,
where $max_{k}$ is the maximum concentration value, $k$, and RefRev is
parameterised as for $\beta_{3}$ in the step length function.

@@

\begin{itemize}
	\item Spatial dynamics: e.g. Figure 14. showing the population movement
		across weeks, including a spawning period.
	\item Overall population trends: e.g. Figure 3, showing the population
		dynamics at play.
	\item Realised step function: e.g. Figure 13, showing how the movement
		and turning angle responds. This could be combined with Figure
		7 for a more complete visualisaiton.
	\item Catch composition and spatial clustering: e.g. Figure 10, showing
		the importance of spatial resolution.
	\item Some measure of temporal trends/changes in spatial composition -
		could pick a certain cell and show how the proportions change
		over time ? An example is Figure 15, though there is little
		variation here. I think this is because the static habitat with
		population movement is far more stable than the previous
		directly affected population distributions (where there was an
		advective-diffusive process from the SPDE). Might need to
		consider how to change this?  Could include a spatiotemporally
		changing habitat suitability covariate, e.g. temperature ??
		\textcolor{red}{[temperature would be very interesting and
			cover varying spatial fields]}
	\item Comparison of population structure from i) commercial sampling,
		ii) fisheries independent survey, iii) real population. This
		should be some statistical measure...not sure best approach
		here.
	\item [If we perform the closure, some measure of cost/benefit based on
		different resolutions of data and data used (i.e. commercial,
		survey, real populations). This could be on the basis of
		closing x cells, giving cell references to the model so they
		cannot be fished. Need to consider whether this is also a
		seasonal closure.]
\end{itemize}
